LAGOS lake depth

### 2019-10-30 lagos_depth.Rmd
classify shape classes

```{r hypsography, eval=TRUE, fig.height = 4}
ids <- factor(c("convex", "concave"))
polys <- data.frame(
  id = rep(ids, each = 3),
  x = c(0, 75, 0, 
        25, 100, 100), 
  y = c(75, 0, 0, 
        100, 100, 25))

gg_hypso_ct <- ggplot(data = hypso_ct) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
  theme(legend.position = "bottom")

@@

gg_hypso_ct + theme(plot.title = element_blank())

@@

```{r raster-to-profile-to-hypso, eval=TRUE, fig.height = 4}
# 3 panel figure of a single starting with rayshader, then depth profile, then hypsography

@@

```{r gg_rayshader, eval=TRUE}
library(rayshader)
library(elevatr)

@@

```{r gg-rayshader-concave, eval=TRUE}

@@

```{r gg-rayshader-convex, eval=TRUE}
res_rsub2 <- prep_rayshader(rsub2, elev2)

@@

```{r get-hypso, eval=TRUE, fig.height = 4}
# "pickerel_lake" 7096 # convex
# "long_pond" 99642 # concave
rsub1 <- raster::raster("data/ct_bathy/7096.tif")
rsub2 <- raster::raster("data/ct_bathy/99642.tif")
names(rsub1) <- "layer"
names(rsub2) <- "layer"

@@

```{r get-hypso-normalized, eval=TRUE, fig.height = 4}
gg_radius_normal <- ggplot(data = hypso, 
                           aes(x = radius_percent, y = depth_percent)) +
    geom_point(aes(color = type)) + geom_line(aes(color = type)) +
    xlab("Radius distance (percent)") + ylab("Depth (percent)") + 
    ylim(c(100, 0)) +  theme(legend.position = "none")

@@

```{r workflow, eval=TRUE}
gg_radius_convex <- 
  ggplot(data = dplyr::filter(hypso, type == "convex"), 
         aes(x = radius_percent, y = depth)) +
    geom_point() + geom_line() +
    xlab("Radius distance (proportion)") + ylab("Depth") +
    scale_y_reverse()

@@

```{r compare-slopes, eval=TRUE, fig.height = 4}
dt <- dplyr::filter(lg$lakes_limno,
                    !is.na(maxdepth) & !is.na(meandepth)) %>%
  dplyr::select(lagoslakeid, nhd_lat, nhd_long, maxdepth, meandepth) %>%
  dplyr::left_join(dplyr::select(lg$locus, lagoslakeid, lake_area_ha, gnis_name, state_zoneid)) %>%
  left_join(dplyr::select(lg$buffer100m.lulc,
                          buffer100m_slope_mean, lagoslakeid)) %>%
  dplyr::left_join(dplyr::select(lg$state, state_zoneid, state)) %>%
  dplyr::filter(lake_area_ha >= 4 & lake_area_ha <= 400) %>%
  dplyr::filter(state == "MN") %>%
  dplyr::filter(meandepth != maxdepth) %>%
  dplyr::filter(meandepth > 1)

@@

```{r eval=TRUE, fig.height = 4}
plot(dt$half_slope, dt$slope)
abline(0, 1)
```

```{r }

>>>>

```{r hypsography, eval=TRUE, fig.height = 3.2}
ids <- factor(c("convex", "concave"))
polys <- data.frame(
  id = rep(ids, each = 3),
  x = c(0, 100, 0, 
        0, 100, 100), 
  y = c(100, 0, 0, 
        100, 100, 0))

gg_hypso_ct <- ggplot(data = hypso_ct) +
  geom_polygon(data = polys, aes(x = x, y = y, group = id, fill = id), alpha = 0.2) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid)) +
  geom_abline(intercept = -100, slope = 1, color = "red") +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
  theme(legend.position = c(0.84, 0.2), 
        legend.title = element_blank())

@@

gg_hypso_ct <- gg_hypso_ct + theme(plot.title = element_blank())
# for poster
ggdraw(gg_hypso_ct, ylim = c(0, 1), clip = "off") + 
  draw_label("200 MN and CT lakes with bathymetry [1,4]", 
             y = 0.06, x = 0.22, size = 8) +
  draw_label("33 percent concave, 66 percent convex", 
             y = 0.01, x = 0.2, size = 8)
```

```{r classify, eval = TRUE}
# round the x value and test if y is greater or less than a straight-line curve
hypso_ct_classes <- hypso_ct %>%
  group_by(WBNAME) %>%
  summarize_all(median) %>%
  mutate(area_percent_rounded = round(area_percent)) %>%
  left_join(data.frame(y = rev(seq(0, 100, by = 1)), x = seq(0, 100, by = 1)), 
            by = c("area_percent_rounded" = "x")) %>%
  data.frame() %>%
  mutate(shape_class = case_when(depth_percent < y ~ "convex", 
                                 depth_percent > y ~ "concave", 
                                 TRUE ~ "neither")) %>%
  ungroup() %>%   dplyr::filter(!is.na(shape_class)) %>% 
    dplyr::select(WBNAME, llid, shape_class)

# table(hypso_ct_classes$shape_class)

hypso_ct_w_class <- left_join(hypso_ct_classes, hypso_ct)

ggplot(data = hypso_ct_w_class) +
  geom_line(aes(x = area_percent, y = depth_percent, group = llid, 
                color = shape_class)) +
  geom_abline(intercept = -100, slope = 1) +
  ylim(c(100, 0)) +
  ylab("Depth (percent)") + xlab("Area (percent)") + 
  ggtitle(paste0("CT", " (", length(unique(hypso_ct$llid)), ") lakes")) + 
  theme(legend.position = "bottom") +
  labs(fill = "")
```

```{r class-depths, eval = TRUE}

distinct(hypso_ct_w_class, llid, .keep_all = TRUE) %>%
  dplyr::filter(shape_class != "neither") %>%
  group_by(shape_class) %>%
  summarize(depth = list(tibble::enframe(
    quantile(depth, probs = c(0.05,0.5,0.95), na.rm = TRUE)))) %>% 
  unnest() %>%
  tidyr::spread(name, value) %>%
  ggplot() +
geom_pointrange(aes(x = shape_class, y = `50%`, 
                      ymin = `5%`, ymax = `95%`)) +
  ylab("Depth (m)") + xlab("")

@@

```{r raster-to-profile-to-hypso, eval=FALSE, fig.height = 4}
# 3 panel figure of a single starting with rayshader, then depth profile, then hypsography

@@

```{r gg_rayshader, eval=FALSE}
library(rayshader)
library(elevatr)

@@

```{r gg-rayshader-concave, eval=FALSE}

@@

```{r gg-rayshader-convex, eval=FALSE}
res_rsub2 <- prep_rayshader(rsub2, elev2)

@@

```{r get-hypso, eval=FALSE, fig.height = 4}
# "pickerel_lake" 7096 # convex
# "long_pond" 99642 # concave
rsub1 <- raster::raster("data/ct_bathy/7096.tif")
rsub2 <- raster::raster("data/ct_bathy/99642.tif")
names(rsub1) <- "layer"
names(rsub2) <- "layer"

@@

```{r get-hypso-normalized, eval=FALSE, fig.height = 4}
gg_radius_normal <- ggplot(data = hypso, 
                           aes(x = radius_percent, y = depth_percent)) +
    geom_point(aes(color = type)) + geom_line(aes(color = type)) +
    xlab("Radius distance (percent)") + ylab("Depth (percent)") + 
    ylim(c(100, 0)) +  theme(legend.position = "none")

@@

```{r workflow, eval=FALSE}
gg_radius_convex <- 
  ggplot(data = dplyr::filter(hypso, type == "convex"), 
         aes(x = radius_percent, y = depth)) +
    geom_point() + geom_line() +
    xlab("Radius distance (proportion)") + ylab("Depth") +
    scale_y_reverse()

@@

```{r compare-slopes, eval=FALSE, fig.height = 4}
dt <- dplyr::filter(lg$lakes_limno,
                    !is.na(maxdepth) & !is.na(meandepth)) %>%
  dplyr::select(lagoslakeid, nhd_lat, nhd_long, maxdepth, meandepth) %>%
  dplyr::left_join(dplyr::select(lg$locus, lagoslakeid, lake_area_ha, gnis_name, state_zoneid)) %>%
  left_join(dplyr::select(lg$buffer100m.lulc,
                          buffer100m_slope_mean, lagoslakeid)) %>%
  dplyr::left_join(dplyr::select(lg$state, state_zoneid, state)) %>%
  dplyr::filter(lake_area_ha >= 4 & lake_area_ha <= 400) %>%
  dplyr::filter(state == "MN") %>%
  dplyr::filter(meandepth != maxdepth) %>%
  dplyr::filter(meandepth > 1)

@@

```{r eval=FALSE, fig.height = 4}
plot(dt$half_slope, dt$slope)
abline(0, 1)
```
